---
title: "ShilpiSharma_Aug_SVAP_Asmt_R2"
author: "Shilpi Sharma"
date: "10/8/2017"
output:
  pdf_document: default
  html_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Domain - Employment (People)
### Topic - Unemployment Analysis and Comparison at the country and gender level

```{r}
# Loading required libraries
library(rvest)
library(tidyr)
library(dplyr)
library(ggplot2)
library(pander)
library(tidyverse)
library(readxl)
library(stringr)
library(RColorBrewer)
library(lattice)
library(reshape2)
library(gridExtra)
```

## Frame the questions
- Which are the top 10 countries with the highest unemployment ratio of males to females for both the timeframes 1991 and 2016?
- Which country has the highest takers of intermediate and advanced education in 2016?


## Acquire the Data
Getting the unemployment data for different countries from the World Bank database 

```{r}
setwd("/Users/shilpisharma/SVAPData")
getwd()
UData=read_excel("Unemployment.xls")
summary(UData)
attach(UData)
str(UData)
```

## Refine the Data
- Check for Quality and Consistency
- Missing values
- Outlier treatment
- Remove unneeded
- Format Data Types

### Changing Column Names to shorter names
```{r}
dim(UData)
column_name <- c('country', 'maleUnempt91', 'maleUnempt16', 'femaleUnempt91', 'femaleUnempt16', 'maleYouthUnempt91', 'maleYouthUnempt16', 'femaleYouthUnempt91', 'femaleYouthUnempt16', 'basicEduPct', 'intermediateEduPct', 'advancedEduPct')
colnames(UData) <- column_name
str(UData)
```

### Change the data types from chr to numeric and date
```{r}
head(UData)
tail(UData)
# Coercing all the columns datatype except Country from chr to numeric -> NAs introduced
UData$maleUnempt91 <- as.numeric(UData$maleUnempt91)
UData$maleUnempt16 <- as.numeric(UData$maleUnempt16)
UData$femaleUnempt91 <- as.numeric(UData$femaleUnempt91)
UData$femaleUnempt16 <- as.numeric(UData$femaleUnempt16)
UData$maleYouthUnempt91 <- as.numeric(UData$maleYouthUnempt91)
UData$maleYouthUnempt16 <- as.numeric(UData$maleYouthUnempt16)
UData$femaleYouthUnempt91 <- as.numeric(UData$femaleYouthUnempt91)
UData$femaleYouthUnempt16 <- as.numeric(UData$femaleYouthUnempt16)
UData$basicEduPct <- as.numeric(UData$basicEduPct)
UData$intermediateEduPct <- as.numeric(UData$intermediateEduPct)
UData$advancedEduPct <- as.numeric(UData$advancedEduPct)
tail(UData)
```

### Filter all the rows except the last four rows 
```{r}
df <- UData %>% filter(row_number() < 223) 
tail(df)
# Removing the special characters in Korea country field
df <- df %>% within(country[str_detect(country, 'Korea, Dem+')] <- 'Korea Dem.')
tail(df)
```

## Transform the Data
### First Question - Finding the top 10 countries with the highest unemployment ratio of males to females
```{r}
# Calculating for 1991 data
uemptRatio91 = df$maleUnempt91/df$femaleUnempt91 
df <- cbind(df, uemptRatio91)
str(df)
dfUemptRatio91 <- df %>% arrange(desc(uemptRatio91)) %>% head(10)

# Calculating for 2016 data
uemptRatio16 = df$maleUnempt16/df$femaleUnempt16 
df <- cbind(df, uemptRatio16)
str(df)
dfUemptRatio91 <- df %>% arrange(desc(uemptRatio91)) %>% head(10)
dfUemptRatio16 <- df %>% arrange(desc(uemptRatio16)) %>% head(10)
```

### Solution : Uganda had the highest unemployment ratio of males to females in 1991 but it moved down to spot 10 in 2016, whereas Sierra has the highest unemployement ratio in 2016.  

## Explore - Visualize 
### Depicting through Bar Graph
```{r}
# Depicting 1991 data through Bar Graph
ggplot(dfUemptRatio91) + 
  aes(reorder(country, uemptRatio91), uemptRatio91, fill=country) + 
  geom_col(width = 1) + xlab("Country") + ylab("Unemployment Ratio - Males to Females") +
  coord_flip()

# Depicting 2016 data through Bar Graph
ggplot(dfUemptRatio16) + 
  aes(reorder(country, uemptRatio16), uemptRatio16, fill=country) + 
  geom_col(width = 1) + xlab("Country") + ylab("Unemployment Ratio - Males to Females") +
  coord_flip()
```

### Depicting through pie charts
```{r}
# Plotting 1991 data on a pie chart
plot1 <- ggplot(dfUemptRatio91, facets = ~mygroup) + 
  aes(reorder(country, uemptRatio91), uemptRatio91, fill=country) + 
  geom_col(width = 1) + xlab("Country") + ylab("Unemployment Ratio - Males to Females (1991)") +
  coord_flip() + 
  coord_polar()  

# Plotting 2016 data on a pie chart
plot2 <- ggplot(dfUemptRatio16, facets = ~mygroup) + 
  aes(reorder(country, uemptRatio16), uemptRatio16, fill=country) + 
  geom_col(width = 1) + xlab("Country") + ylab("Unemployment Ratio - Males to Females (2016)") +
  coord_flip() + 
  coord_polar()

grid.arrange(plot1, plot2, ncol=2)
```

### Depicting both the 1991 and 2016 unemployment ratio data through same Scatter Plot 
```{r}
# Depicting 1991 data through Scatter Plot 
ggplot(dfUemptRatio91) + 
  aes(country, uemptRatio91) + 
  geom_point()

# Depicting 2016 data through Scatter Plot 
ggplot(dfUemptRatio16) + 
  aes(country, uemptRatio16) + 
  geom_point()

# Depicting both timeframes in the same scatter plot
# Combining the 1991 and 2016 ratio data frames
d <- rbind(dfUemptRatio91, dfUemptRatio16)
# Removing duplicate rows
d <- unique(d)
# Extracting the columns of interest
d1 <- as.data.frame(cbind(country=d$country, U2016=d$uemptRatio16, U1991=d$uemptRatio91))
str(d1)
# Transposing the data from wide to long format
df2 <- melt(data = d1, id = "country")
# Renaming the variable column name to Year
colnames(df2)[colnames(df2) == 'variable'] <- 'Year'
# Plotting on a scatter plot
ggplot(data = df2, aes(x = country, y = value, colour = Year, group = 1)) + geom_point() + xlab("Country") + ylab("Unemployment Ratio of Males to Females") + scale_x_discrete(labels = abbreviate) 
```

## Second Question - Which country has the highest takers of intermediate and advanced education in 2016?
### Stacked Bar Plot with Colors and Legend
```{r}
# Extracting the required education indicator columns from the cleansed data frame
percents <- as.data.frame(cbind(country = df$country, intermediateEducation = df$intermediateEduPct, advancedEducation = df$advancedEduPct))
# Removing the NAs
percents <- subset(percents, !is.na(intermediateEducation) & !is.na(advancedEducation))
# Getting the top 10 intermediate education takers
topIntermediate <- percents %>% arrange(desc(intermediateEducation)) %>% head(10)
# Getting the top 10 advanced education takers
topAdvanced <- percents %>% arrange(desc(advancedEducation)) %>% head(10)
# Combining the Intermediate top ten and Advanced top ten data
combinedEduData <- unique(rbind(topIntermediate, topAdvanced))
str(combinedEduData)
summary(combinedEduData)
# Depicting through stacked bar chart
# Transposing the data from wide to long format
df3 <- melt(data = combinedEduData, id = "country")
# Renaming the variable column name to Education Type
colnames(df3)[colnames(df3) == 'variable'] <- 'EducationType'
ggplot(df3, aes(x=country)) + geom_bar(aes(fill = EducationType)) + xlab("Country") + 
  ylab("Percentage Distribution of Intermediate and Advanced education takers") + coord_flip()
```

### Conclusion - Insights gained
* Uganda has the highest disproportion between males and females with males being more unemployed in 1991 timeframe. 
* Uganda reduced this ratio from 2.16 to 1.42 in 2016.
* Sierra Leone had the highest unemployment ratio of males to females in 2016.
* Brazil has highest number of people completing intermediate education.
* Croatia has the highest number of people completing the advanced education.

